Numerical trajectory patterns of floating macroalgae

### 2020-11-05 thesis_chapter_3.Rmd
changes to script and tables updated

p_RK4 <- tidync("OceanParcels/trajectory_data/PassiveFloatRK4.nc") %>%
  hyper_tibble() %>%
  dplyr::select(lon, lat, time, distance, age) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2002-01-01 00:00:00"))

p_RK4_mean <- p_RK4 %>% # summarise data based on time 
   group_by(t) %>% 
   summarise_all(funs(mean = mean, sd = sd, min = min, max = max), na.rm = T)

@@

p_EE <- tidync("OceanParcels/trajectory_data/PassiveFloatEE.nc") %>%
  hyper_tibble() %>%
  dplyr::select(lon, lat, time, distance, age) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2002-01-01 00:00:00"))

p_EE_mean <- p_EE %>% # summarise data based on time 
   group_by(t) %>% 
   summarise_all(funs(mean = mean, sd = sd, min = min, max = max), na.rm = T)

p_EE$type <- "EE" # add extra column 'type' to dataframe and assigned 'EE' label

k_app1 <- tidync("OceanParcels/trajectory_data/KelpFloatApp1.nc") %>%
  hyper_tibble() %>%
  dplyr::select(lon, lat, time, distance, age) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2002-01-01 00:00:00"))

k_app1_mean <- k_app1 %>% # summarise data based on time 
   group_by(t) %>% 
   summarise_all(funs(mean = mean, sd = sd, min = min, max = max), na.rm = T)

k_app1$type <- "Approach 1" # add extra column 'type' to dataframe and assigned 'kelp' label

k_app2 <- tidync("OceanParcels/trajectory_data/KelpFloatApp2.nc") %>%
  hyper_tibble() %>%
  dplyr::select(lon, lat, time, distance, age) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2002-01-01 00:00:00"))

k_app2_mean <- k_app2 %>% # summarise data based on time 
   group_by(t) %>% 
   summarise_all(funs(mean = mean, sd = sd, min = min, max = max), na.rm = T)

k_app2$type <- "Approach 2" # add extra column 'type' to dataframe and assigned 'kelp' label

total_data <- rbind(p_RK4, p_EE, k_app1, k_app2)

@@

heatmap_EE <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
  geom_hex(data = p_EE, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  guides(fill=guide_legend(title="Number of particles")) +
  theme_bw()

heatmap_APP1 <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  #geom_bin2d(data = k_app1, bins = 80, aes(x = lon, y = lat)) +
  geom_hex(data = k_app1, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  guides(fill=guide_legend(title="Number of particles")) +
  theme_bw()

heatmap_APP2 <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
  geom_hex(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  guides(fill=guide_legend(title="Number of particles")) +
  theme_bw()

@@

library(ggpubr)
ggarrange(heatmap_RK4, heatmap_EE, heatmap_APP1, heatmap_APP2, labels = c("A","B","C","D"), common.legend = TRUE, legend = c("right"), ncol = 2, nrow = 2)

@@

```{r scatter plot processing, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}

@@

```{r age plots, warning=FALSE, echo=FALSE, include=TRUE, message=FALSE}

@@

```{r morphology_wrangling, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
# Investigate morphology data to assist in estimating surface area
# Load software
library(tidyverse)
library(ggpubr)
library(gridExtra)
library(ggplot2)
library(vegan)
library(dplyr)
# Read in morphology data
eck_df <- read.csv("data/morph_data/eck_morph_data.csv", sep = ";") %>%
  filter(site!= "Rocky Bank") %>%   # Not using Rocky Bank in this study
  filter(site!= "De Hoop") 
eck_df$site <- as.character(eck_df$site)
eck_df <- eck_df %>% 
  filter(population == "Deep") 
eck_long_df <- gather(eck_df, Morphology, measurement, frond_mass:total_length, factor_key=TRUE)
# Summary stats to make it easier to refer to later, these will not be used any analysis.
site_list <- c("Kommetjie")
eck_stats <- eck_long_df %>% 
  group_by(site, Morphology) %>% 
  summarise(mean = mean(measurement, na.rm = T),
            minimum = min(measurement, na.rm = T),
            maximum = max(measurement, na.rm = T)) %>% 
  ungroup() %>% 
  filter(site %in% site_list) %>% 
  rename(Site = site) %>% 
  rename(Mean = mean) %>% 
  rename(Minimum = minimum) %>% 
  rename(Maximum = maximum)

>>>>

p_RK4 <- tidync("OceanParcels/trajectory_data/RK4_PFloat.nc") %>%
  hyper_tibble() %>%
  dplyr::select(lon, lat, time, distance, age) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))

@@

Mean_Kfloat_H100_W00 <- tidync("OceanParcels/trajectory_data/Mean_KFloat_H100_W00.nc") %>%
  hyper_tibble() %>%
  dplyr::select(lon, lat, time, distance, age) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))

Mean_Kfloat_H100_W00$type <- "Mean_Kfloat_H100_W00" # add extra column 'type' to dataframe and assigned 'kelp' label

Mean_KFloat_H095_W05 <- tidync("OceanParcels/trajectory_data/Mean_KFloat_H095_W05.nc") %>%
  hyper_tibble() %>%
  dplyr::select(lon, lat, time, distance, age) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))

Mean_KFloat_H095_W05$type <- "Mean_KFloat_H095_W05" # add extra column 'type' to dataframe and assigned 'kelp' label

Mean_KFloat_H090_W10 <- tidync("OceanParcels/trajectory_data/Mean_KFloat_H090_W10.nc") %>%
  hyper_tibble() %>%
  dplyr::select(lon, lat, time, distance, age) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))

Mean_KFloat_H090_W10$type <- "Mean_KFloat_H090_W10" # add extra column 'type' to dataframe and assigned 'kelp' label

Mean_KFloat_H085_W15 <- tidync("OceanParcels/trajectory_data/Mean_KFloat_H085_W15.nc") %>%
  hyper_tibble() %>%
  dplyr::select(lon, lat, time, distance, age) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2018-01-01 00:00:00"))

Mean_KFloat_H085_W15$type <- "Mean_KFloat_H090_W10" # add extra column 'type' to dataframe and assigned 'kelp' label

total_data <- rbind(p_RK4, Mean_Kfloat_H100_W00, Mean_KFloat_H095_W05, Mean_KFloat_H090_W10, Mean_KFloat_H085_W15)

@@

heatmap_Mean_Kfloat_H100_W00 <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
  geom_hex(data = Mean_Kfloat_H100_W00, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  guides(fill=guide_legend(title="Number of particles")) +
  theme_bw()

heatmap_Mean_KFloat_H095_W05 <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  #geom_bin2d(data = k_app1, bins = 80, aes(x = lon, y = lat)) +
  geom_hex(data = Mean_KFloat_H095_W05, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  guides(fill=guide_legend(title="Number of particles")) +
  theme_bw()

heatmap_Mean_KFloat_H090_W10 <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
  geom_hex(data = Mean_KFloat_H090_W10, bins = 80, aes(x = lon, y = lat)) +
  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  guides(fill=guide_legend(title="Number of particles")) +
  theme_bw()

heatmap_Mean_KFloat_H085_W15 <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  #geom_bin2d(data = k_app2, bins = 80, aes(x = lon, y = lat)) +
  geom_hex(data = Mean_KFloat_H085_W15, bins = 80, aes(x = lon, y = lat)) +
  scale_fill_continuous(type = "viridis", limits = c(1, 13000)) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  guides(fill=guide_legend(title="Number of particles")) +
  theme_bw()

@@

library(ggpubr)
ggarrange(heatmap_RK4, heatmap_Mean_Kfloat_H100_W00, heatmap_Mean_KFloat_H095_W05, heatmap_Mean_KFloat_H090_W10, heatmap_Mean_KFloat_H085_W15, labels = c("A","B","C","D", "E"), common.legend = TRUE, legend = c("right"), ncol = 2, nrow = 3)

@@

```{r scatter plot processing, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}

@@

```{r age plots, warning=FALSE, echo=FALSE, include=TRUE, message=FALSE, eval=FALSE}

@@

```{r morphology_wrangling, eval=TRUE, message=FALSE, warning=FALSE, include=TRUE}
# Investigate morphology data to assist in estimating surface area
# Load software
library(tidyverse)
library(ggpubr)
library(gridExtra)
library(ggplot2)
library(vegan)
library(dplyr)
# Read in morphology data
eck_df <- read.csv("data/morph_data/eck_morph_data.csv", sep = ";") %>%
  filter(site!= "Rocky Bank") %>%   # Not using Rocky Bank in this study
  filter(site!= "De Hoop") 
eck_df$site <- as.character(eck_df$site)
eck_df <- eck_df %>% 
  filter(population == "Deep") 
eck_long_df <- gather(eck_df, Morphology, measurement, frond_mass:total_length, factor_key=TRUE)
# Summary stats to make it easier to refer to later, these will not be used any analysis.
site_list <- c("Kommetjie")
eck_stats <- eck_long_df %>% 
  group_by(site, Morphology) %>% 
  summarise(mean = mean(measurement, na.rm = T),
            minimum = min(measurement, na.rm = T),
            maximum = max(measurement, na.rm = T)) %>% 
  ungroup() %>% 
  filter(site %in% site_list) %>% 
  rename(Site = site) %>% 
  rename(Mean = mean) %>% 
  rename(Minimum = minimum) %>% 
  rename(Maximum = maximum)
